---
title: "AIMS.MSRC - exporting intensity matrix of IMS data"
output:
  html_document: default
  html_notebook: default
---

# Introduction
Often times after processing data we would like to export the peak table to share with collaborators or use in other software or even use in other R packages. In this brief tutorial we show how to do this. We will use the ```exportIntensMatrix()``` function for this purpose. This function writes a comma-separated table to disk that can easily be loaded into Microsoft Excel or other spreadsheet program or other software of the intensity data for every pixel and mass bin of a Cardinal dataset.

In this context, metadata are things like sample name, any ROIs selected through ```selectROI()```, and anything else in the ```pData(cardinaldata)```. 

### Import required packages
```{r import_packages, echo=T, message=F, warning=FALSE, results="hide"}
require(Cardinal)
require(aimsMSRC)
```

### Import included mouse brain cerebellum dataset
```{r load_data}
data(cer_dan_peaks)
```

### Exporting the peak table with all metadata over the full mass range
```{r outmeta1}
fullMZ_fullMeta <- exportIntensMatrix(cer_dan_peaks, 
                                  mz_range = c(0,100000), #we give this a much larger range than our actual dataset 
                                  metadata = TRUE, 
                                  filename = "fullMZ_fullMeta") # this will be appended with .csv extension at write-time
```

### Exporting the peak table without metadata over the full mass range
```{r outmeta2}
fullMZ_noMeta <- exportIntensMatrix(cer_dan_peaks, 
                                  mz_range = c(0,100000), #we give this a much larger range than our actual dataset 
                                  metadata = FALSE,
                                  filename = "fullMZ_noMeta")
```

### Exporting the peak table with all metadata over select mass range
```{r outmeta3}
mz725_750_fullMeta <- exportIntensMatrix(cer_dan_peaks, 
                                  mz_range = c(700,750), #we only export a peak table with between m/z 700-750
                                  metadata = TRUE, 
                                  filename = "mz700_750_fullMeta")
```

We can get an idea of the structure of these R data.frames by looking at the first few columns and rows as shown below:

```{r outmeta1_tbl}
head(fullMZ_fullMeta, row.names=F)[,1:6] # the last indexing only prints the first 6 columns
```

```{r outmeta2_tbl}
head(fullMZ_noMeta, row.names=F)[,1:6] # the last indexing only prints the first 6 columns
```

```{r outmeta3_tbl}
head(mz725_750_fullMeta , row.names=F)[,1:6] # the last indexing only prints the first 6 columns
```

*Now the data we have processed and examined in R can be used in other softwares or other R packages.*
