---
title: "AIMS.MSRC - IMS data pre-processing"
output:
  html_document: default
  html_notebook: default
---

# Introduction

In this tutorial we will load an .imzML converted imaging mass spectrometry dataset into Cardinal and pre-process it in the following ways:

1. Normalization to the total ion current (TIC)
2. Peak picking
3. Peak alginment
4. Peak filtering

In many instances, imaging mass spectrometry data acquired by TOF mass spectrometers will need additional pre-processing in the form of signal smoothing and baseline correction. We won't demo these with this lipid dataset, but their use is thoroughly explained in the [Cardinal documentation.](http://bioconductor.org/packages/release/bioc/vignettes/Cardinal/inst/doc/Cardinal-walkthrough.pdf)

### Import required packages
```{r import_packages, echo=T, message=F, warning=FALSE, results="hide"}
require(Cardinal)
```

### Read imzML file into Cardinal
```{r read_data}
cer_dan <- readImzML('DAN_cerebellum_neg', folder = 'D:/code/testing_data/') # data file : D:/code/testing_data/DAN_cerebellum_neg.imzML

```

```{r ds_info_print, echo=F}
cat('Dataset info:\n',
    'm/z bins \n',
    '\t',dim(cer_dan)[[1]],'\n',
    'no. pixels \n' ,
    '\t',dim(cer_dan)[[2]])
```

### TIC normalization
```{r normalize_data, message=FALSE, warning=FALSE, results="hide"}
cer_dan <- normalize(cer_dan, method = "tic")
```

### Initial peak picking
```{r peak_pick_init, message=FALSE, warning=FALSE, results="hide"}
cer_dan_pl <- peakPick(cer_dan, 
                   pixel = seq(1, ncol(cer_dan), by = 10), #we peak pick on every 10th spectrum here
                   method = "simple", #simple assumes constant noise accross mass axis
                   SNR = 3) #SNR = signal to noise, 3 is standard but this will vary by dataset
```

### Peak alignment to mean
```{r peak_align, message=FALSE, warning=FALSE, results="hide"}
cer_dan_pl_aligned <-  peakAlign(cer_dan_pl, #unaligned peak list
                         ref = cer_dan, #generate mean spec
                         method = "diff", #search by difference
                         units = "mz", #search by m/z
                         diff.max = 0.35) #plus minus m/z tolerance

```

### Peak filtering
```{r peak_filter, message=FALSE, warning=FALSE, results="hide"}
cer_dan_pl_filtered <- peakFilter(cer_dan_pl_aligned, #aligned peaklist
                                        method = "freq", # by peak frequency
                                        freq.min = ncol(cer_dan_pl_aligned)*0.05) #peak must be in at least 5% of the spectra

```

```{r peak_info_print, echo=F}
cat('retained peaks \n',
    '\t',dim(cer_dan_pl_filtered)[[1]])
```

### Peak extraction from whole dataset
```{r peak_extraction, eval=F, message=FALSE, warning=FALSE, results="hide"}
##this process can be time consuming
cer_dan_peaks <- reduceDimension(cer_dan, #normalized full profile dataset
                                 ref = cer_dan_pl_filtered, #developed peak list
                                 type = "height") #get peak heights (intensities)
```

*Now we have a peak picked dataset in the Cardinal container that we can use for data analysis.*
We can save this dataset to disk and do other data analysis steps now.

### Saving RData for later use
```{r, eval = F}
save(cer_dan_peaks, file="cer_dan_peaks.RData")
```

### Loading RData for later use
```{r, eval = F}
load("cer_dan_peaks.RData")
```
