---
title: "AIMS.MSRC - ROI comparison within a single sample"
output:
  html_document: default
  html_notebook: default
---

# Introduction

Within this example we will show how to do region of interest comparison within a single sample and some algorithms that are useful for making such comparisons.

### Import required packages
```{r import_packages, echo=T, message=F, warning=FALSE, results="hide"}
require(Cardinal)
require(aimsMSRC)
```

### Import included mouse brain cerebellum dataset
```{r}
data(cer_dan_peaks)
```

### Plot ion image
```{r}
image(cer_dan_peaks, mz=885.58)
```

### Select an ROI based on a displayed ion image
*n.b. this command must be run in RStudio in order to make selection and is not compatible with notebook*
```{r selectROI, eval=F}
##molecular layer
cer_dan_peaks = selectROI(cer_dan_peaks, mz=766.55)
```

```{r show_molecular}
image(cer_dan_peaks, cer_dan_peaks$molecular_layer ~ x*y, 
      strip=F, 
      colorkey=F, 
      key=F,
      main="Molecular layer selection in red")
```
```{r, eval=F}
##granular layer
cer_dan_peaks = selectROI(cer_dan_peaks, mz=795.58)
```

```{r show_granular}
image(cer_dan_peaks, cer_dan_peaks$granular_layer ~ x*y, 
      strip=F, 
      colorkey=F, 
      key=F,
      main="Granular layer selection in red")
```

### Reciever Operator Characteristic (ROC) curve analysis to determine discriminant signals
*molecular layer*
```{r ROC_molecular}
##determine the species most associated with the molecular layer
mol_layer_ROC <- rocIMS(cer_dan_peaks, 
       mzs= 'all', 
       positiveROI = 'molecular_layer',
       negativeROI = 'all',
       undersampling=T)

##print top 6 m/zs of molecular layer by ROC analysis:
cat(head(names(mol_layer_ROC)))

```

```{r ROC_plot_molecular}
plotROCmz(cer_dan_peaks, mol_layer_ROC, top_n = 3, layout=c(2,3))
```

*granular layer*
```{r ROC_granular}
##determine the species most associated with the molecular layer
gran_layer_ROC <- rocIMS(cer_dan_peaks, 
       mzs= 'all', 
       positiveROI = 'granular_layer',
       negativeROI = 'all',
       undersampling=T)

##print top 6 m/zs of molecular layer by ROC analysis:
cat(head(names(gran_layer_ROC)))

```

```{r ROC_plot_granular}
plotROCmz(cer_dan_peaks, gran_layer_ROC, top_n = 3, layout=c(2,3))
```

### Look at intensity distribution of these ROIs
```{r int_boxplot}
mzIntensPlot(cer_dan_peaks, mz=795.589, roi_names = c('molecular_layer','granular_layer'))

mzIntensPlot(cer_dan_peaks, mz=834.518, roi_names = c('molecular_layer','granular_layer'))

```

